### 2025-02-26 paper/performance.tex
PyTorch numbers.

One of the goals of this work is to demonstrate that it is possible to
formulate the problem in a proof assistant and then pass it on to the
other system that can run the algorithm efficiently. In order to
substantiate this claim, we compare the running times of the code that
we generate from the specification of the CNN at the end of the
Section~\ref{sec:edsl} with an equivalent CNN implemented with
TensorFlow~\cite{ad-tf}. This is a limited study, and it does not
exploit all the expressivity provided by our language, but
nevertheless it shows the potential of achieving performance that
approaches that of established programs in non-verified languages. We
use the commonly used MNIST database, which consists of 60000
greyscale images of handwritten digits, each 28 by 28
pixels~\cite{deng2012mnist}.

@@

Our experimental results have been obtained on an NVIDIA A100 with
CUDA 11.8 and cuDNN 8.6.0, and are shown in
Table~\ref{tab:performance}. The runtime numbers exclude startup
costs, such as graph compilation, initialising GPU context, copying
data to the GPU, in order to focus on the performance of the generated
code. For the largest training set, these startup costs are about $5s$
seconds for TensorFlow and $6.7s$ for Futhark, which dwarfs the actual
training time, due to our small datasets.

@@

\begin{table}
\begin{tabular}{crrr}
\textbf{Size of training set} & \multicolumn{2}{c}{\textbf{Runtime}} & \textbf{Ratio} \\
& Futhark & TensorFlow & \\
$10000$ & $0.91s$ & $1.07s$ & $0.85\times{}$ \\
$60000$ & $4.93s$ & $2.92s$ & $1.68\times{}$
\end{tabular}
\caption{Training time for various training set sizes, comparing
  Futhark and Tensorflow, running on an A100 GPU. The third column
  shows the speedup of TensorFlow compared to Futhark. We use a batch
  size of $1000$, a training rate of $0.05$, and train for $10$
  epochs.}
\label{tab:performance}
\end{table}

>>>>

One of the goals of this work is to demonstrate that it is possible to
formulate the problem in a proof assistant and then pass it on to the
other system that can run the algorithm efficiently. In order to
substantiate this claim, we compare the running times of the code that
we generate from the specification of the CNN at the end of the
Section~\ref{sec:edsl} with an equivalent CNN implemented with
TensorFlow~\cite{ad-tf} and PyTorch~\cite{ad-pytorch}. This is a
limited study, and it does not exploit all the expressivity provided
by our language, but nevertheless it shows the potential of achieving
performance that approaches that of established programs in
non-verified languages. We use the commonly used MNIST database, which
consists of 60000 greyscale images of handwritten digits, each 28 by
28 pixels~\cite{deng2012mnist}.

@@

Our experimental results have been obtained on an NVIDIA A100 with
CUDA 11.8 and cuDNN 8.6.0, and are shown in
Table~\ref{tab:performance}. The runtime numbers exclude startup
costs, such as graph compilation, initialising GPU context, copying
data to the GPU, in order to focus on the performance of the generated
code. For the largest training set, these startup costs are about $5s$
seconds for TensorFlow and $6.7s$ for Futhark, which dwarfs the actual
training time, due to our small datasets. The PyTorch runtime is
extremely high, due to low GPU utilization.

@@

\begin{table}
\begin{tabular}{crrrr}
\textbf{Size of training set} & \multicolumn{3}{c}{\textbf{Runtime}} & \textbf{Ratio} \\
& Futhark & TensorFlow & PyTorch & \\
$10000$ & $0.91s$ & $1.07s$ & $19s$ & $0.85\times{}$ \\
$60000$ & $4.93s$ & $2.92s$ & $61s$ & $1.68\times{}$
\end{tabular}
\caption{Training time for various training set sizes, comparing
  Futhark, Tensorflow, and PyTorch, running on an A100 GPU. The final column
  shows the speedup of TensorFlow compared to Futhark. We use a batch
  size of $1000$, a training rate of $0.05$, and train for $10$
  epochs.}
\label{tab:performance}
\end{table}
